To be FAIR: Theory Specification Needs an Update

### 2025-06-07 manuscript.tex
update 2025-06-07 15:14:35.120732

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\makeatother
\keywords{fairtheory, meta science, theory formation, open science\newline\indent Word count: 9491}
\usepackage{lineno}

@@

This is a preprint paper, generated from Git Commit \# 5d7ad1a4. The first and senior author positions were fixed, and remaining author positions randomized using \texttt{set.seed(6999982);\ sample(c("Aaron",\ "Felix",\ "Noah",\ "Maximilian"))}.

@@

We propose FAIR theory as an instantiation of modular publishing (Kircz, 1998).
In most fields, the primary unit of scientific communication is the academic paper.
A paper may depend on multiple resources - materials, data, code, and theory - but these are often merely described in the text.
Modular publishing is the practice of making each of these resources available as independent citable \emph{information artifacts},
with adequate metadata that is indexed in standardized repositories (Van De Sompel et al., 2004).
Data sharing is an example of a modular publishing practice that is widely adopted and increasingly required by funding agencies, journals, and universities.
Scholars can archive information artifacts in repositories like \href{https://zenodo.org/}{Zenodo},
which was developed by \href{https://home.cern/}{CERN} under the European Union's \href{https://www.openaire.eu/}{OpenAIRE} program (European Organization For Nuclear Research \& OpenAIRE, 2013).
To maintain a persistent record of scholarly communication,
Zenodo mints DOIs for information artifacts - as does, for example, the \href{https://www.crossref.org/}{Crossref} association,
which is used by many academic publishers.
Finally, the DataCite Metadata Schema offers a standard way to document the nature of relationships between information artifacts (DataCite Metadata Working Group, 2024).
For example, a dataset collected for a specific paper would be archived in Zenodo with the metadata property \texttt{resourceType:\ dataset},
and cross-reference the published paper with \texttt{relationType:\ IsSupplementTo}.
Similarly, a FAIR theory object can be connected to a specific paper which might serve as the theory's documentation and canonical reference by using \texttt{relationType:\ IsDescribedBy}, while the reverse relationship, documented in the canonical reference paper, is \texttt{relationType:\ Describes}. Other types are useful for making relationships between multiple theory objects explicit: If an existing theory is made FAIR without substantial alterations,
the resulting FAIR theory metadata would cross-reference the existing theory as \texttt{relationType:\ IsDerivedFrom}.
If an existing theory is updated, \texttt{relationType:IsNewVersionOf} could be used to reference previous versions.
If a variation of an existing FAIR theory is created, cross-reference it with \texttt{relationType:\ IsVariantFormOf}.

@@

If we consider the current state of Reusability in psychological theory, there appears to be a norm against theory reuse:
\emph{``{[}Theories are{]} like toothbrushes --- no self-respecting person wants to use anyone else's''} (Mischel, 2008).
As cumulative knowledge acquisition requires reusable theories that are continuously updated based on insights from new data, such a norm impedes scientific progress (De Groot \& Spiekerman, 1969).
In FAIR theory workshops, we similarly notice a reluctance to reuse and adapt existing theories.
Students ask questions such as ``Who owns a theory?'',
and ``Who determines how a theory may be reused or changed?''.
These questions imply a norm against modifying theory without its author's consent, reminiscent of the aforementioned problem of dependency on the author.

Licensing theories for reuse unambiguously answers these questions,
with the caveats that legislation may vary across contexts and jurisdictions, and that this paper does not constitute legal advice.
Two considerations are important when determining what license is appropriate for theory.
A first consideration is that copyright law protects authors' rights according to the idea-expression dichotomy (Bently et al., 2010).
It explicitly does not
\emph{``extend to any idea, procedure, process, system, method of operation, concept, principle, or discovery''}.
Copyright thus applies to creative works expressing a theory (e.g., prose, visual illustrations),
but not to the underlying theoretical idea.
It thus seems that theories expressed in prose or depicted visually - in other words, that fall short of the Accessibility criterion - are more likely to qualify for copyright protection than formal theories.
A second consideration is that academic research is covered under ``fair use'' exemptions to copyright.
Given these two considerations - that copyright does not protect ideas in their purest form and that academic use offers exemptions to copyright - it may be counterproductive and possibly misleading to adopt a license that assumes copyright protection to theories.
For psychological theories without commercial aspects, we suggest using a licence that explicitly waives copyright and encourages Reusability,
such as CC0 (no rights reserved).

@@

Several authors have reinterpreted De Groot's empirical cycle.
An important advantage of FAIR theory is that we can implement different versions of a theory, compare them, and document their cross-relationships.
We can take work that has been done before - in this case, the repository created above, and create an independent copy that we can modify as we wish, while retaining cross-references to the original.
For example, the DOT graph below implements Wagenmakers and colleagues' (2018) interpretation of the empirical cycle.
First, notice that the phases of the cycle have been renamed.
This change was not described in the paper.
If we assume that the labels are meant to illustrate the phases,
not substantially change the ontology,
then we can represent it by adding labels to the original DOT graph.
These labels suggest a focus on empirical psychology that was not present in De Groot's version.
Furthermore, the label ``knowledge accumulation'' invites the question of exactly \emph{how} knowledge accumulates upon evaluation of a prior experiment.
As this lack of cumulative knowledge acquisition appears to be precisely where contemporary research practice falls short, this ambiguity invites further improvement of the theory.
The authors explicitly mention a second change: \emph{``We added the Whewell-Peirce-Reichenbach distinction between the context of discovery and the context of justification''}.
The DOT graph below shows our implementation of this version of the empirical cycle, by adding subgraphs.

@@

The first author was inspired by De Groot as well,
but again specified the empirical cycle differently.
First, in De Groot's formulation, each stage describes a process.
This invites the question of what the concrete outcomes of these processes are.
In other words: what actually changes when going through the cycle, except the scholar's mind?
To address this point, the nodes in Van Lissa's specification refer to specific deliverables, whereas the edges connecting the nodes refer to processes acting upon those deliverables, see Figure \ref{fig:figecs}c).
Second, the processes of induction and deduction are perhaps not as neatly confined to specific phases as De Groot proposed.
Theory testing, as takes place in the ``context of justification'', can be said to involve mostly deductive reasoning.
Theory development and amendment, as takes place in the ``context of discovery'', involves primarily inductive reasoning\footnote{Here, ``induction'' is defined forming general theories based on specific observations. Others have used the term ``abduction'' to describe ``inference to the best explanation'' (Peirce, 1960). For present purposes, the terms are interchangeable.}.
However, deriving hypotheses from theory is not purely deductive
as auxiliary assumptions must often be made to account for remaining ambiguities in theory, which involves induction.
A rudimentary example is assuming equal variances across groups when testing a mean difference between groups, because groups often have equal variances.
Similarly, if we consider the claim that observation is theory-laden, then it too involves induction (Brewer \& Lambert, 2001).
Furthermore, if the testing procedure is not explicitly defined before seeing the data, it incurs some inductive bias as well (Peikert, 2023).
These alterations result in the following implementation of the empirical cycle:

@@

FAIR theory requires a departure from contemporary practice.
Several factors can expedite such a culture change.
One key factor is the \emph{recognition and rewards} movement:
practices for evaluating scientific output are evolving, with initiatives like the \href{https://sfdora.org/read/}{\emph{Declaration on Research Assessment} (DORA)} and \href{https://coara.eu/}{Coalition for Advancing Research Assessment}
promoting the use of more diverse and meaningful metrics beyond journal impact factors.
Modular publishing capitalizes on these changing metrics,
and publishing theories as citeable artifacts allows scholars to be credited for contributions to theory (Kircz, 1998).
Journals that publish theoretical papers could require authors to additionally publish their theories in a FAIR format, cross-referencing the paper,
to expedite its effective reuse and iterative improvement.
A second factor is to lower barriers to adopting FAIR theory by building upon existing widely adopted open science infrastructures.
For this reason, we advocate the use of Git for version control, Zenodo for archiving, and DataCite for standardized metadata.
Barriers to entry can also be lowered by simplifying workflows, which is the goal of the \texttt{theorytools} R-package.
Fourth, the availability of Open Educational Materials (OEM) about theory development contributes to doctoral socialization.
These materials allow teachers to incorporate theory development into their curriculum without investing substantial time on course development,
thus educating the next generation to make use of and contribute to FAIR theory.
Finally, community building plays an important role;
the international network of open science communities, reproducibility networks, and other similar initiatives provide platforms for disseminating FAIR theories and related methodological innovations.
Authors can also share their FAIR theories with other early adopters by submitting them to the ``FAIR Theory Community'' on Zenodo.

@@

\phantomsection\label{fairtheoryprereg}{One important future direction is embedding FAIR theories withing existing open science methodologies.
For example, consider how FAIR theory relates to preregistration.
These practices are clearly distinct but complimentary.
The purpose of FAIR theory is to communicate general principles and expectations about a given phenomenon,
and to provide infrastructure for explicitly deriving hypotheses from specific theories and revising those theories in light of empirical results.
The purpose of preregistration, by contrast, is to eliminate inductive bias from hypothesis tests and increase trust in their outcomes (Peikert et al., 2023).
These practices differ in their level of abstraction:
FAIR theories cut across studies on a given phenomenon,
whereas preregistrations are specific implementations of hypothesis tests,
with a specific study design, analysis plan, and - optionally - a fully reproducible analysis pipeline.
These practices complement each other:
authors can make the derivation chain from theory to hypothesis more explicit by citing a specific FAIR theory in their preregistration.
Moreover, it is possible to preregister an inferential procedure that would require revising the theory after observing data,
or even to have proponents and detractors of a theory review a registered report of such a test.
In short, combining FAIR theory with preregistration and other existing open science practices has the potential to strengthen the epistemic cycle of prediction, testing, and revision, moving us closer to a cumulative science.}

Another future direction is the intersection between the aforementioned ``theory crisis'' and the related ``measurement crisis'' pertaining to the lack of clarity, consistency, and validity in the operationalization of theoretical constructs (Bringmann et al., 2022).
Since FAIR theories can reference other theories and resources,
it is possible to attach references to specific measurement instruments (or even theories of measurement) to constructs named in a theory.
\phantomsection\label{jinglejangle}{FAIR theories can also help address ``jingle- and jangle fallacies'' in psychology,
which are ambiguities that arise from using the same term for different constructs, or conversely, using different terms for the same construct (\textbf{songLiteratureReviewingAddressing2021?}).
By explicitly referencing operational definitions in FAIR theories,
such jingle-jangle fallacies would come to light and could ultimately be resolved.}

@@

FAIR theory is a major step forward towards more transparent, collaborative, and efficient theory construction.
It provides much-needed open science methods for the inductive phase of the empirical cycle,
closing a critical gap in the scientific process.
FAIR theory makes theory more tangible, enabling scholars to incorporate it in their day-to-day workflows in order to derive hypotheses, select control variables, and contribute new data-driven insights.
This paves the way for more theory-driven scholarship
and accelerates cumulative knowledge acquisition in psychology, the social sciences, and beyond.

@@

\bibitem[\citeproctext]{ref-nosekPromotingOpenResearch2015a}
Nosek, B. A., Alter, G., Banks, G. C., Borsboom, D., Bowman, S. D., Breckler, S. J., Buck, S., Chambers, C. D., Chin, G., Christensen, G., Contestabile, M., Dafoe, A., Eich, E., Freese, J., Glennerster, R., Goroff, D., Green, D. P., Hesse, B., Humphreys, M., \ldots{} Yarkoni, T. (2015). Promoting an open research culture. \emph{Science}, \emph{348}(6242), 1422--1425. \url{https://doi.org/10.1126/science.aab2374}

@@

@@

@@

\bibitem[\citeproctext]{ref-vanlissaTeacherCornerEvaluating2020}
Van Lissa, C. J., Gu, X., Mulder, J., Rosseel, Y., Zundert, C. V., \& Hoijtink, H. (2020). Teacher's {Corner}: {Evaluating Informative Hypotheses Using} the {Bayes Factor} in {Structural Equation Models}. \emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{0}(0), 1--10. \url{https://doi.org/10.1080/10705511.2020.1745644}

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\makeatother
\keywords{fairtheory, meta science, theory formation, open science\newline\indent Word count: 9794}
\usepackage{lineno}

@@

We are particularly endebted to our Reviewer, Dr.~Eric Turkheimer, whose thoughtful reflections on FAIR theory were incorporated into the Discussion. We also thank the Editor, Dr.~Scott Vrieze, and two anonymous Reviewers for their detailed and invaluable comments.

@@

We envision FAIR theory as an instance of modular publishing (Kircz, 1998).
\phantomsection\label{modularvalue}{The primary unit of scientific communication has long been the academic paper.
In the process of writing papers, scholars often produce many valuable resources, including instruments, materials, data, code, and theory.
These resources are often merely described in papers (if at all), and not made available for reuse.
Modular publishing is the practice of making each of these resources available as independent \emph{information artifacts},
facilitating their reuse and making them citable (Van De Sompel et al., 2004).}
\phantomsection\label{datasharing}{Data sharing is an example of modular publishing that is already widely adopted, becoming the de-facto standard in psychology (Tedersoo et al., 2021).}

Modular publishing can be achieved by archiving specific resources (including theory) in repositories like \href{https://zenodo.org/}{Zenodo},
which was developed by \href{https://home.cern/}{CERN} under the European Union's \href{https://www.openaire.eu/}{OpenAIRE} program (European Organization For Nuclear Research \& OpenAIRE, 2013).
To maintain a persistent record of scholarly communication,
Zenodo mints DOIs for information artifacts.
The DataCite Metadata Schema offers a standard way to document information artifacts with relevant metadata which increases their findability and documents their relationships to other resources (DataCite Metadata Working Group, 2024).
For example, data can be archived in Zenodo with the metadata property \texttt{resourceType:\ dataset}.
If the data were collected for a specific paper, that relationship can be cross-referenced with \texttt{relationType:\ IsSupplementTo}.
Similarly, a FAIR theory can be connected to a specific paper by using \texttt{relationType:\ IsDescribedBy}, while the reverse relationship, documented in the canonical reference paper, is \texttt{relationType:\ Describes}.
Other cross-references are useful for documenting relationships between multiple theory objects: If an existing theory is made FAIR without substantial alterations,
the resulting FAIR theory metadata would cross-reference the existing theory as \texttt{relationType:\ IsDerivedFrom}.
If an existing theory is updated, \texttt{relationType:IsNewVersionOf} could be used to reference previous versions.
If a variation of an existing FAIR theory is created, cross-reference it with \texttt{relationType:\ IsVariantFormOf}.
\phantomsection\label{modularpublisingdetract}{Modular publishing of resources, including theories, increases their reuse potential and makes them citable without detracting from the conventional academic paper as a unit of academic communication which allows for argumentation,
nuance,
and the author's voice.
}

@@

If we consider the current state of Reusability in psychological theory, there appears to be a norm against theory reuse:
\emph{``{[}Theories are{]} like toothbrushes --- no self-respecting person wants to use anyone else's''} (Mischel, 2008, p. 1).
As cumulative knowledge acquisition requires reusable theories that are continuously updated based on insights from new data, such a norm impedes scientific progress (De Groot \& Spiekerman, 1969).
In FAIR theory workshops, we similarly notice a reluctance to reuse and adapt existing theories.
Students ask questions such as ``Who owns a theory?'',
and ``Who determines how a theory may be reused or changed?''.
These questions imply a norm against modifying theory without its author's consent, reminiscent of the aforementioned problem of dependency on the author.

Licensing theories for reuse unambiguously answers these questions,
with the caveats that legislation may vary across contexts and jurisdictions, and that this paper does not constitute legal advice.
Two considerations are important when determining what license is appropriate for theory.
A first consideration is that copyright law protects authors' rights according to the idea-expression dichotomy (Bently et al., 2010).
Copyright does not
\emph{``extend to any idea, procedure, process, system, method of operation, concept, principle, or discovery''} (Section 102, Copyright Act).
Copyright thus applies to creative works expressing a theory (e.g., prose, visual illustrations),
but not to the underlying theoretical idea.
It thus seems that theories expressed in prose or depicted visually - in other words, that fall short of the Accessibility criterion - are more likely to qualify for copyright protection than formal theories.
A second consideration is that academic research is covered under ``fair use'' exemptions to copyright.
Given these two considerations - that copyright does not protect ideas in their purest form and that academic use offers exemptions to copyright - it may be counterproductive and possibly misleading to adopt a license that assumes copyright protection to theories.
For psychological theories without commercial aspects, we suggest using a licence that explicitly waives copyright and encourages Reusability,
such as CC0 (no rights reserved).

@@

Several authors have reinterpreted De Groot's empirical cycle.
An important advantage of FAIR theory is that we can implement different versions of a theory, compare them, and document their cross-relationships.
We can take work that has been done before - in this case, the repository created above, and create an independent copy that we can modify as we wish, while retaining cross-references to the original.
For example, the DOT graph below implements Wagenmakers and colleagues' (2018) interpretation of the empirical cycle.
First, notice that the phases of the cycle have been renamed.
This change was not described in the paper.
If we assume that the labels are meant to illustrate the phases,
not substantially change the ontology,
then we can represent it by adding labels to the original DOT graph.
These labels suggest a focus on empirical psychology that was not present in De Groot's version.
Furthermore, the label ``knowledge accumulation'' invites the question of exactly \emph{how} knowledge accumulates upon evaluation of a prior experiment.
As this lack of cumulative knowledge acquisition appears to be precisely where contemporary research practice falls short, this ambiguity invites further improvement of the theory.
The authors explicitly mention a second change: \emph{``We added the Whewell-Peirce-Reichenbach distinction between the context of discovery and the context of justification''} (p.~423).
The DOT graph below shows our implementation of this version of the empirical cycle, by adding subgraphs.

@@

The first author was inspired by De Groot as well,
but again specified the empirical cycle differently.
First, in De Groot's formulation, each stage describes a process.
This invites the question of what the concrete outcomes of these processes are.
In other words: what actually changes when going through the cycle, except the scholar's mind?
To address this point, the nodes in Van Lissa's specification (2025) refer to specific deliverables, whereas the edges connecting the nodes refer to processes acting upon those deliverables, see Figure \ref{fig:figecs}c).
Second, the processes of induction and deduction are perhaps not as neatly confined to specific phases as De Groot proposed.
Theory testing, as takes place in the ``context of justification'', can be said to involve mostly deductive reasoning.
Theory development and amendment, as takes place in the ``context of discovery'', involves primarily inductive reasoning\footnote{Here, ``induction'' is defined forming general theories based on specific observations. Others have used the term ``abduction'' to describe ``inference to the best explanation'' (Peirce, 1960). For present purposes, the terms are interchangeable.}.
However, deriving hypotheses from theory is not purely deductive
as auxiliary assumptions must often be made to account for remaining ambiguities in theory, which involves induction.
A rudimentary example is assuming equal variances across groups when testing a mean difference between groups, because groups often have equal variances.
Similarly, if we consider the claim that observation is theory-laden, then it too involves induction (Brewer \& Lambert, 2001).
Furthermore, if the testing procedure is not explicitly defined before seeing the data, it incurs some inductive bias as well (Peikert, 2023).
These alterations result in the following implementation of the empirical cycle:

@@

FAIR theory requires a departure from contemporary practice.
Several factors can expedite such a culture change.
One key factor is the \emph{recognition and rewards} movement:
practices for evaluating scientific output are evolving, with initiatives like the \href{https://sfdora.org/read/}{\emph{Declaration on Research Assessment} (DORA)} and \href{https://coara.eu/}{Coalition for Advancing Research Assessment}
promoting the use of more diverse and meaningful metrics beyond journal impact factors.
Modular publishing capitalizes on these changing metrics,
and FAIR theory allows scholars to be credited for theoretical contributions (Kircz, 1998).
Journals that publish theoretical papers could require authors to additionally publish their theories in a FAIR format, cross-referencing the paper,
to expedite its effective reuse and iterative improvement.
A second factor is to lower barriers to adopting FAIR theory by building upon existing widely adopted open science infrastructures.
For this reason, we advocate the use of Git for version control, Zenodo for archiving, and DataCite for standardized metadata.
Barriers to entry can also be lowered by simplifying workflows, which is the goal of the \texttt{theorytools} R-package.
Fourth, the availability of Open Educational Materials (OEM) about theory development contributes to doctoral socialization.
These materials allow teachers to incorporate theory development into their curriculum without investing substantial time on course development,
thus educating the next generation to make use of and contribute to FAIR theory.
Finally, community building plays an important role;
the international network of open science communities, reproducibility networks, and other similar initiatives provide platforms for disseminating FAIR theories and related methodological innovations.
Authors can also share their FAIR theories with other early adopters by submitting them to the ``FAIR Theory Community'' on Zenodo.

@@

\phantomsection\label{fairtheoryprereg}{One important future direction is embedding FAIR theories withing existing open science methodologies.
For example, consider how FAIR theory relates to preregistration.
These practices are distinct but complimentary.
The purpose of FAIR theory is to communicate general principles and expectations about a given phenomenon,
and to provide infrastructure for explicitly deriving hypotheses from specific theories and revising those theories in light of empirical results.
The purpose of preregistration, by contrast, is to eliminate inductive bias from hypothesis tests and increase trust in the outcomes of a specific empirical study (Peikert et al., 2023).
FAIR theories are specified at a level of abstraction that transcends individual studies.
FAIR theories can inform - and be informed by - both quantitative and qualitative research.
Preregistrations, by contrast, are specific implementations of quantitative hypothesis tests,
within the context of a specific study design, analysis plan, and - optionally - a fully reproducible analysis pipeline.
These practices complement each other:
authors can make the derivation chain from theory to hypothesis more explicit by citing a specific FAIR theory in their preregistration.
Moreover, it is possible to preregister an inferential procedure that would require revising the theory after observing data,
or even to have proponents and detractors of a theory review a registered report of such a test.
In short, combining FAIR theory with preregistration and other existing open science practices has the potential to strengthen the epistemic cycle of prediction, testing, and revision, moving us closer to a cumulative science.}

\phantomsection\label{measurementcrisis}{Another future direction is the intersection between the aforementioned ``theory crisis'' and the related ``measurement crisis'' pertaining to the lack of clarity, consistency, and validity in the operationalization of theoretical constructs (Bringmann et al., 2022).
Since FAIR theories can reference other theories and resources,
it is possible to attach references to specific measurement instruments (or even theories of measurement) to constructs named in a theory.}
\phantomsection\label{jinglejangle}{FAIR theories can also help address ``jingle- and jangle fallacies'' in psychology,
which are ambiguities that arise from using the same term for different constructs, or conversely, using different terms for the same construct (Song et al., 2021).
By explicitly referencing operational definitions in FAIR theories,
such jingle-jangle fallacies would come to light and could ultimately be resolved.}

@@

FAIR theory is a major step forward towards more transparent, collaborative, and efficient theory construction.
It provides much-needed open science methods for the inductive phase of the empirical cycle,
closing a critical gap in the scientific process.

\phantomsection\label{turkheimer}{We invite the reader to ask the rhetorical question raised by our reviewer, prof. Turkheimer, who - reflecting on a long career as a practitioner of theory in psychology - considered ``how adopting a FAIR framework would change, improve, or inhibit what I do''.
We paraphrase his answer here\footnote{Edited for brevity.}, because his lived experience complements our youthful ``techno-enthusiasm''.
Prof.~Turkheimer remarked that a FAIR framework would have been very helpful in maintaining meaningful boundaries between a theory and the predictions it makes,
in tracking how empirical findings result in changes to empirical claims made by a theory,
and in resolving theoretical disputes.
\emph{``FAIR theory provides a structured, machine- and human-readable framework that distills and formalizes the core components of a theory.
Everything published about a theory can continue to exist in its original form,
while a FAIR representation supplements that literature by explicitly outlining the theory's assumptions, logical structure, and testable implications.
Importantly, FAIR theory can evolve semi-independently from individual papers, offering a persistent and collaborative object that others can reuse, cite, and refine.
Contributing to the development of a FAIR theory thus becomes a new way of contributing to the theoretical literature - much like writing a high-quality review article, but with the added benefit of interoperability, versioning, and reusability.''}}

We envision a future where application of the FAIR principles makes theories more \emph{useful}, living up to Herbert Simon's ideal,
enabling scholars to incorporate theory into their workflows in a tanglible way,
providing explicit derivation chains for hypotheses,
applying transparent rules to select the right control variables for causal inference,
and proposing specific changes to existing theories based on empirical results.
This paves the way for more theory-driven scholarship
and accelerates cumulative knowledge acquisition in psychology, the social sciences, and beyond.

@@

\bibitem[\citeproctext]{ref-nosekPromotingOpenResearch2015}
Nosek, B. A., Alter, G., Banks, G. C., Borsboom, D., Bowman, S. D., Breckler, S. J., Buck, S., Chambers, C. D., Chin, G., Christensen, G., Contestabile, M., Dafoe, A., Eich, E., Freese, J., Glennerster, R., Goroff, D., Green, D. P., Hesse, B., Humphreys, M., \ldots{} Yarkoni, T. (2015). Promoting an open research culture. \emph{Science}, \emph{348}(6242), 1422--1425. \url{https://doi.org/10.1126/science.aab2374}

@@

\bibitem[\citeproctext]{ref-songLiteratureReviewingAddressing2021}
Song, Y., Watson, R. T., \& Zhao, X. (2021). \emph{Literature {Reviewing}: {Addressing} the {Jingle} and {Jangle Fallacies} and {Jungle Conundrum Using Graph Theory} and {NLP}}. Forty-{Second International Conference} on {Information Systems}.

@@

\bibitem[\citeproctext]{ref-tedersooDataSharingPractices2021}
Tedersoo, L., Küngas, R., Oras, E., Köster, K., Eenmaa, H., Leijen, Ä., Pedaste, M., Raju, M., Astapova, A., Lukner, H., Kogermann, K., \& Sepp, T. (2021). Data sharing practices and data availability upon request differ across scientific disciplines. \emph{Scientific Data}, \emph{8}(1), 192. \url{https://doi.org/10.1038/s41597-021-00981-0}

@@

\bibitem[\citeproctext]{ref-vanlissaTeachersCornerEvaluating2020}
Van Lissa, C. J., Gu, X., Mulder, J., Rosseel, Y., Zundert, C. V., \& Hoijtink, H. (2020). Teacher's {Corner}: {Evaluating Informative Hypotheses Using} the {Bayes Factor} in {Structural Equation Models}. \emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{0}(0), 1--10. \url{https://doi.org/10.1080/10705511.2020.1745644}
